---
title: "Cleaning and Kaplein Meyer"
author: "Kaitlin Rommelfanger"
date: "2026-03-26"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(readr)
library(tidyverse)

nursery<- read_csv("Container Experiment Myra/32626CleaingCheck.csv")

#checking values to see that there is only expected levels
unique(nursery$Treatment)
unique(nursery$`Dead/Alive`)
unique(nursery$`Monitoring Session`)


nursery$`Dead/Alive`<-as.factor(nursery$`Dead/Alive`)
nursery$`Epicotyl length (cm)`<-as.numeric(nursery$`Epicotyl length (cm)`)

#look at number alive over time
survivorship<-nursery%>%
  group_by(Treatment, `Dead/Alive`,`Monitoring Session`)%>%
  summarise(count=n())%>%
  pivot_wider(names_from = `Monitoring Session`, values_from = count)


#look at number of plants monitored in each montoring session 
moncheck<-nursery%>%
  group_by(`Monitoring Session`,`Propagule ID`)%>%
  summarise(n=n())%>%#total number of times monitored at each monitoring session
  filter(n>1)#filter for ones monitored more than once in a single monitoring session

unique(moncheck$`Propagule ID`)
#657 665 722 635 607 619 627 680 688

#there were still some trees monitored  twice in the same monitoring sesh, edited in csv and redownloaded, will recheck 

nursery<- read_csv("Container Experiment Myra/32626checkcleaned.csv")

#look at number of plants monitored in each montoring session 
moncheck<-nursery%>%
  group_by(`Monitoring Session`,`Propagule ID`)%>%
  summarise(n=n())%>%#total number of times monitored at each monitoring session
  filter(n>1)#filter for ones monitored more than once in a single monitoring session

#no more duplicates, but there is quite a few trees missing tag IDs now in a few sessions (the number of NAs in propagule ID)
#2=3, 4=13, 11=3, 22=5,25=31,32=50

#sessions 25 and 31 concern me

```


Correcting for trees that were ID ed with the wrong treatment
```{r}

#going to look at number for each treatment montiored in each monitoring session

tcheck<-nursery%>%
  filter(!is.na(`Propagule ID`))%>%
  group_by(`Monitoring Session`, Treatment)%>%
  summarise(N=n())%>%
  pivot_wider(names_from = `Monitoring Session`, values_from = N)

#when I look at the number of trees in each treatment they aren't consistent, sometimes they are over the number of trees planted. want to check to see if trees are consistently marked with the correct treatment

treatmentcheck<-nursery%>%
  group_by(`Propagule ID`)%>%
  summarise(nTreatments=n_distinct(Treatment))%>%
  filter(nTreatments>1)
#we've got 31 trees that are recorded with having more than one treatment, going to generate a list so i can fix it
unique(treatmentcheck$`Propagule ID`)
#[1] 566=TP 568=CC 570=BB 577=TP 582=CC 586 590 598 620 622 627 636 640 642 643 645 646 651 652 653 657 661 665 668 674 681 690 691 714 715 722 725 727 728 

#filtering for a single prop ID so I can look at it over time, see if it was just a typo
check<-nursery%>%
  filter(`Propagule ID`==582)


#going to fix the treatments so that same prop ID has the same treatment the whole time (I added a new column, treatment fixed, not writing over original column)
nursery_treatmentchange <- nursery %>%
  group_by(`Propagule ID`) %>% #for each propagule id
  mutate(
    Treatment_fixed = ifelse(  
      is.na(`Propagule ID`),    #if prop ID is NA want to keep original treatment
      Treatment,
      names(sort(table(Treatment), decreasing = TRUE))[1])) %>% #if not NA do this
  ungroup()

#table(treatment) makes a table with the different treatments and the number of times it is recorderd as being that treatment, (50 times BB, 1 time CC) then it sorts the table values in decending order most to least instances of that treatment, then for each prop ID it replaces the treatment with the most common treatment found, which fixes the times when the trees are randomly IDed with the wrong treatment, so now we have a new column Treatment_fixed that should have all trees only with one treatment over time


#checking how many trees in each treatment with fixed treatment column, we don't have a way to know if the trees without prop IDs have the correct treatment, so we can't include them for that time point

tcheck<-nursery_treatmentchange%>%
  filter(!is.na(`Propagule ID`))%>% #filtering out NAs
  group_by(`Monitoring Session`, Treatment_fixed)%>%
  summarise(N=n())%>%
  pivot_wider(names_from = `Monitoring Session`, values_from = N)


#I have 10 time points with 60 trees monitored in each treatment, a lot of the other monitoring points are missing trees in treatments. 

```

looking at living dead over time
```{r}
#using this to visualize survivorship over time for individual trees and see when we have missing sessions
survive<-nursery_treatmentchange%>%
  select(`Monitoring Session`, `Propagule ID`, Treatment, `Dead/Alive`)%>%
  pivot_wider(names_from = `Monitoring Session` , values_from =  `Dead/Alive`)

#this is organizing the dataframe by propagule ID and monitoring session then grouping by propagule id, creating a new column called zombie. In zombie the funciton lag() will look at the value  of Dead/Aive at the previous monitoring session and put that value in zombie plants column. then I am filtering for times where dead/alive is A and the previous session trees (my mutated ZombiePlants column) were marked dead
zombies<-nursery_treatmentchange%>%
  arrange(`Propagule ID`, `Monitoring Session`)%>% 
  group_by(`Propagule ID`)%>%
  mutate(ZombiePlants=lag(`Dead/Alive`))%>%
  ungroup()%>%
  filter(`Dead/Alive`=="A")%>%
  filter(ZombiePlants=="D") 

#looking at how many trees there are that are zombies
length(unique(zombies$`Propagule ID`))
#68 trees that are dead then alive again

#what trees are they
unique(zombies$`Propagule ID`)
#560 561 564 566 567 575 580 581 584 585 591 595 603 604 605 606 609 617 622 626 627

#628 634 636 637 639 640 642 645 648 651 652 653 655 657 663 665 667 669 672 674 675 

#677 686 687 688 690 692 694 696 697 698 699 700 702 703 706 707 708 715 716 721 722 724 726 728 731


#at this point I went one by one filtering for the trees and looking at the data, once I identified the time point of "death" I wrote over all points either moving forward as D or backward as alive using and if else statement

#I also added resprout column and populated that with yes in the case I found resprouted trees

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="560")
#560 marking dead at first death session 3looks like it was decaped, no notes of resprout

#so I am going to rewerite over the D/A column marking dead after first death unless in the case of the resprout trees
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="560"&`Monitoring Session`>=3,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="560")
#if else statement works!! Boom shacka lacka now we do them all  


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="561")
#561 dead after first dead, session 5, decap no notes of resprout
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="561"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="564")
#RESPROUT, marking dead after session 3
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="564"&`Monitoring Session`>=3,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="564",
    "Yes",
    NA
  ))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="566")
#gonna call it dead after 2 but noting RESPROUT decap noted, later sprouting noted but epi lengths gone after RESPROUT
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="566"&`Monitoring Session`>=2,
         "D",
         `Dead/Alive`))
#now adding resprout column
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="566",
    "Yes",
    Resprout
  ))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="567")
#mark dead after session 4, no resprout
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="567"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="575")
#mark dead session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="575"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="580")
#dead session 6
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="580"&`Monitoring Session`>=6,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="581")
#dead after first dead session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="581"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="584")
#decap but resprout dead after 7 RESPROUT
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="584"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="584",
    "Yes",
    Resprout
  ))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="585")
#dead at first monitoring session
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="585"&`Monitoring Session`>=1,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="591")
#mark dead at session 4 and all following
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="591"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="595")
#dead at session 7
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="595"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="603")
#dead at session 7
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="603"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="604")
#dead at session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="604"&`Monitoring Session`>=2,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="605")
#dead session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="605"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="606")
#not dead at all
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="606"&`Monitoring Session`>=1,
         "A",
         `Dead/Alive`))



check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="609")
#dead at session 7
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="609"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="617")
#dead at session 3
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="617"&`Monitoring Session`>=3,
         "D",
         `Dead/Alive`))



check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="622")
#dead at session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="622"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="626")
#dead at session 3
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="626"&`Monitoring Session`>=3,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="627")
#dead at session 3
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="627"&`Monitoring Session`>=3,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="628")
#dead at session 3
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="628"&`Monitoring Session`>=3,
         "D",
         `Dead/Alive`))



check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="634")
#dead at session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="634"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="636")
#dead at session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="636"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="637")
#dead at session 9
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="637"&`Monitoring Session`>=9,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="637"&`Monitoring Session`<9,
         "A",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="639")
#alive for all sessions!
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="639"&`Monitoring Session`>=1,
         "A",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="640")
#dead at session 2
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="640"&`Monitoring Session`>=2,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="642")
#dead at session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="642"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="645")
#dead at session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="645"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="648")
#dead at session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="648"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="651")
#decap but resprouted Alive whole time (makring dead at 9)
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="651"&`Monitoring Session`>=9,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="651",
    "Yes",
    Resprout
  ))



check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="652")
#broken then resprouted alive the whole time
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="652",
    "Yes",
    Resprout
  ))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="652"&`Monitoring Session`==4,
         "A",
         `Dead/Alive`))



check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="653")
#dead at session 2
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="653"&`Monitoring Session`>=2,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="655")
#dead at session 6
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="655"&`Monitoring Session`>=6,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="655"&`Monitoring Session`<6,
         "A",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="657")
#dead at session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="657"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="657"&`Monitoring Session`<5,
         "A",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="663")
#dead at session 6
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="663"&`Monitoring Session`>=6,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="665")
#decap at 7, new shoot at last monitoring session RESHOOT
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="665"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="665",
    "Yes",
    Resprout
  ))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="667")
#DECAP, RESPROUT happened dead at session 7
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="667"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="667",
    "Yes",
    Resprout
  ))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="669")
#dead at session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="669"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="672")
#dead at session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="672"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="674")
#dead at session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="674"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="675")
#alive all along
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="675"&`Monitoring Session`>=1,
         "A",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="677")
#dead session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="677"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))



#677 686 687 688 690 692 694 696 697 698 699 700 702 703 706 707 708 715 716 721 722 724 726 728 731
check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="686")
#dead at session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="686"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="687")
#dead at session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="687"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="688")
#dead at session 7
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="688"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="690")
#dead session 6
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="690"&`Monitoring Session`>=6,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="692")
#dead at session 6
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="692"&`Monitoring Session`>=6,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="694")
#dead at session 7
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="694"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="696")
#dead session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="696"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="697")
#dead at 6 RESPROUT 
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="697"&`Monitoring Session`>=6,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="697",
    "Yes",
    Resprout
  ))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="698")
#dead session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="698"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="699")
#dead session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="699"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="700")
#dead session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="700"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="702")
#dead session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="702"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="703")
#dead at session 7
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="703"&`Monitoring Session`>=7,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="706")
#dead at session 8
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="706"&`Monitoring Session`>=8,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="707")
#dead session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="707"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="708")
#dead session 5
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="708"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="715")
#dead session 4
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="715"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="716")
#dead at 4th monitoring session (not monitored 2, or 3rd)
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="716"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="721")
#dead at 5th monitoring session 
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="721"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="722")
#dead at 5th monitoring session 
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="722"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="724")
#dead at 5th mon sesh 
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="724"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))


#724 726 728 731
check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="726")
#alive the whole time
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="726"&`Monitoring Session`>=1,
         "A",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="728")
#dead at 4th mon sesh
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="728"&`Monitoring Session`>=4,
         "D",
         `Dead/Alive`))

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="731")
#RESPROUT
nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(`Dead/Alive`=ifelse(
    `Propagule ID`=="731"&`Monitoring Session`>=5,
         "D",
         `Dead/Alive`))

nursery_treatmentchange<-nursery_treatmentchange%>%
  mutate(Resprout=ifelse(
    `Propagule ID`=="731",
    "Yes",
    Resprout
  ))


```


```{r}
#this was a check I ran after adjusting for all zombies, I fixed the trees that were still zombies 
zombiecheck<-nursery_treatmentchange%>%
  arrange(`Propagule ID`, `Monitoring Session`)%>% 
  group_by(`Propagule ID`)%>%
  mutate(ZombiePlants=lag(`Dead/Alive`))%>%
  ungroup()%>%
  filter(`Dead/Alive`=="A")%>%
  filter(ZombiePlants=="D") 

#double check these in the AM

unique(zombiecheck$`Propagule ID`)
#564 604 637 652 655 657


check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="564")

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="604")

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="652")
#a resprout
check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="655")
#fixed 
check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="657")

#when i recheck zombies after fixing this the only remaining one is a resprout woo all zombies slayed

```


myra and I checking our resprots
```{r}
nursery<-nursery_treatmentchange%>%
  group_by(Treatment_fixed, `Dead/Alive`)%>%
  summarise(N=n())


nurserycheck<-nursery_treatmentchange%>%
  filter(Resprout=="Yes")#%>%
  filter(`Monitoring Session`==16)%>%
  group_by(Treatment_fixed)%>%
  summarise(N=n())



nurserycheck<-nursery_treatmentchange%>%
  filter(Resprout=="Yes")%>%
  filter(Treatment_fixed=="CC")

nurserycheck<-

unique(nurserycheck$`Propagule ID`)
#667 652 665 697 731 564 584 566 651

check<-nursery_treatmentchange%>%
  filter(`Propagule ID`=="697")

```




Ok Kaplein Meyer Time!

The theory behind it all
https://biostatsquid.com/survival-time-analysis-easily-explained/
https://biostatsquid.com/kaplan-meier-curve/
https://biostatsquid.com/easy-log-rank-test/
https://biostatsquid.com/easy-cox-regression-for-survival-analysis/

the code (how to do it)
https://biostatsquid.com/easy-survival-analysis-r-tutorial/ 

another tutorial, disclaimer I like the first one better https://www.sthda.com/english/wiki/survival-analysis-basics


```{r}
library(lubridate)#for date things

#in order to run Kaplein Meyer we need to setup the data,

#we need censoring status as 0 still alive and 1 died

#and we need time to event, number of days until D, or last followup, in our case session 32 in order to calculate time until death

#want to look at dates for each monitoring session,make date processed a date
nursery_treatmentchange$`Date Processed`<-mdy(nursery_treatmentchange$`Date Processed`)

#now I'm picking one date for each monitoring session, because sessions run across multiple days, going with the earliest date that trees were monitored
session_dates <- nursery_treatmentchange %>%
  group_by(`Monitoring Session`) %>%
  summarise(
    first_date = min(`Date Processed`, na.rm = TRUE)
  ) %>%
  ungroup()

```



Need to rerun after filtering out 
```{r}
library(survival)
library(ggsurvfit)
library(survminer)

#I'm creating a dataframe here for the test

resproutcheck<-nursery_treatmentchange%>%
  filter(Resprout=="Yes")

unique(resproutcheck$`Propagule ID`)
#9 trees resprouted, could remove these or make it so resprouted ones are A, at the end, but then I'm not sure cuz how do we know that the ones we thought were dead didnt eventually resprout?


#for this plot removed resprouted trees

#here I am looking at the dataframe and if there are any D recorded for the trees, I am selecting for the earliest session where they were dead (makind first_dead_session), otherwise marking as session 32 
status<-nursery_treatmentchange%>%
  filter(is.na(Resprout))%>% #removing resprouted trees
  arrange(`Propagule ID`, `Monitoring Session`) %>%
  group_by(`Propagule ID`) %>%
  summarise(
    first_dead_session = ifelse(
      any(`Dead/Alive` == "D"), #if there are any D
      min(`Monitoring Session`[`Dead/Alive` == "D"]), #then select the lowest dead 
      32  #otherwise fill fist_dead_session column with 32 (survived all sessions)
    ), status = ifelse(  #make status
      any(`Dead/Alive` == "D"), #then if there are any D in D/A
      1,  #mark D as 1
      0) , #if no D, A = 0
    Treatment = first(Treatment_fixed))%>%  #treatment from fixed treatment column
  ungroup()%>%
  filter(!is.na(`Propagule ID`))#removing NAs in propagule ID
  
#now renaming so i can join the dates with the status frame
status<-status%>%
  rename(`Monitoring Session`=first_dead_session)

#merging 
status<-merge(session_dates, status, by="Monitoring Session")

#get time until death/end. using september 16th as the planting day
status<-status%>%
  mutate(time=first_date-as.Date("2021-09-16"))

#make Kaplan-Meier Curve from the survival object
  #survival object is using our data 
curve <- survfit(Surv(time, status) ~ Treatment, data = status)

#rough plot
plot(curve)

#pretty plot!
plotobj<-ggsurvplot(curve, data=status,
  palette = c("olivedrab","cornflowerblue","darkblue"), conf.int = TRUE, 
  risk.table = TRUE,
  risk.table.height=0.25)

plotobj


#testing for statistical difference between groups 
logRank<-survdiff(Surv(time = time, event = status) ~ Treatment, data = status)

logRank

#Call:
#survdiff(formula = Surv(time = time, event = status) ~ Treatment, 
#    data = status)

#              N Observed Expected (O-E)^2/E (O-E)^2/V
#Treatment=BB 59       24     21.6     0.278     0.479
#Treatment=CC 58       11     22.1     5.606     9.799
#Treatment=TP 54       26     17.3     4.364     6.784

# Chisq= 11.4  on 2 degrees of freedom, p= 0.003 

#difference is detected from significant p value 

```


Here I wanted to fill in D/A for the trees that have missing mon seshes for the survivorship plot but I stopped for the update
```{r}
#want to check how many trees have a missed monitoring session
survivalcheck<-nursery_treatmentchange%>%
  select(`Propagule ID`, Treatment_fixed, `Monitoring Session`,`Dead/Alive`)%>%
  pivot_wider(names_from = `Monitoring Session`, values_from = `Dead/Alive`)

#now I want to fill in survivorship for monitoring sessions where the trees were missed. This should add a row fully populated with NAs for each prop ID with missing monitoring sessions 

NurseryComplete <- nursery_treatmentchange %>%
  group_by(`Propagule ID`) %>%
  complete(`Monitoring Session` = full_seq(`Monitoring Session`, 1)) %>% #the 1 means fill sessions in increments of 1
  ungroup()

#checking to see if instead of nulls shouws NAs now
survivalcheck<-NurseryComplete%>%
  select(`Propagule ID`, `Monitoring Session`,`Dead/Alive`)%>%
  pivot_wider(names_from = `Monitoring Session`, values_from = `Dead/Alive`)
#everything looks good, but treatments are now NA  want to fix that along with DeadAlive

library(dplyr)

NurseryComplete1 <- NurseryComplete %>%
  group_by(`Propagule ID`) %>%
  arrange(`Monitoring Session`) %>%
  mutate(
    # Find first session where tree is dead
    first_dead_session = min(`Monitoring Session`[`Dead/Alive` == "D"], na.rm = TRUE),
    
    # If no death exists, set to Inf (so everything stays Alive)
    first_dead_session = ifelse(is.infinite(first_dead_session), Inf, first_dead_session),
    
    `Dead/Alive` = case_when(
      # Keep observed values
      !is.na(`Dead/Alive`) ~ `Dead/Alive`,
      
      # Before first death → Alive
      `Monitoring Session` < first_dead_session ~ "A",
      
      # At or after death → Dead
      `Monitoring Session` >= first_dead_session ~ "D"
    )
  ) %>%
  ungroup()

#check
survivalcheck<-NurseryComplete1%>%
  select(`Propagule ID`, `Monitoring Session`,`Dead/Alive`)%>%
  pivot_wider(names_from = `Monitoring Session`, values_from = `Dead/Alive`)

#there are a lot of trees missing from session 32, i think i might pull them for the plot

#because the added rows are all NAs now we have Nas in treatment, needs to be fixed

NurseryComplete1 <- NurseryComplete1 %>%
  group_by(`Propagule ID`) %>% #for each propagule id
  mutate(
    Treatment_fixed = ifelse(  
      is.na(`Propagule ID`),    #if prop ID is NA want to keep original treatment
      Treatment,
      names(sort(table(Treatment), decreasing = TRUE))[1])) %>% #if not NA do this
  ungroup()

```



```{r}

SurvivalPlot<-NurseryComplete1%>%
  filter(`Monitoring Session`!=32)%>% #removing 32 so many missing
  filter(!is.na(`Propagule ID`))%>% #removing prop NAs, these survival values filled
  group_by(`Monitoring Session`, Treatment_fixed,`Dead/Alive`)%>%
  summarise(N=n())%>%
  ggplot(aes(x=`Monitoring Session`, y=N, fill=`Dead/Alive`))+
  scale_fill_manual(labels=c("Alive", "Dead"), values=(c("darkgreen", "darkorange")))+
  geom_col(position="stack")+
  facet_wrap(~Treatment_fixed)

SurvivalPlot

#im still seeing more than 60 trees in a treatment, problem, its becasue at first I was still including trees that had NAs for propagule ID, once filtered those out looked good! Also filtered out monitoring session 32
```

Looking at height at a couple time points mean and standard deviation across treatments so i can estimate effect size for my power analysis


```{r}
effectsizethings<-NurseryComplete%>%
  group_by(Treatment, `Monitoring Session`)%>%
  filter(!is.na(`Epicotyl length (cm)`))%>%
  summarise(AVG=mean(`Epicotyl length (cm)`),SD=sd(`Epicotyl length (cm)`))



NurseryTest<-NurseryComplete%>%
  filter(!is.na(`Epicotyl length (cm)`))


#I'm picking one date for each monitoring session, because sessions run across multiple days, going with the earliest date that trees were monitored
session_dates <- NurseryTest %>%
  group_by(`Monitoring Session`) %>%
  summarise(
    first_date = min(`Date Processed`, na.rm = TRUE)
  ) %>%
  ungroup()

NurseryTest$`Monitoring Session`
session_dates$`Monitoring Session`

#join on mon sesh date
NurseryTest<-NurseryTest%>%
  left_join(session_dates, by="Monitoring Session")


#now mutate subtracting starting date
NurseryTest<-NurseryTest%>%
  mutate(Age=first_date-as.Date("2021-09-24"))

NurseryTest<-NurseryTest%>%
  filter(!is.na(`Propagule ID`))%>%
  filter(!is.na(Age))%>%
  filter(!is.na(`Epicotyl length (cm)`))

NurseryTest$Age_years <- NurseryTest$Age / 365.25

#trying to determine effect size of different treatments to get an idea of treatment effect for power analyis 
library(lme4)
model<-lmer(`Epicotyl length (cm)`~Treatment_fixed*Age_years+(Age_years|`Propagule ID`), data=NurseryTest)


model

library(effectsize)

eta <- eta_squared(model, partial = TRUE)
eta

```

